Towards semantic-rich word embeddings

### 2019-03-12 paper/main.tex
Remove [H] from figures due to weird formatting
Improve our method section

%% Useful packages
%\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{tabu}
\usepackage{array}

@@

\section{Our method}
\subsection{Keyword embedding}
\label{sec:kw_embed}
\textit{Keyword} is a sequence of words that lets us disambiguate between different meanings of the same ambiguous word, e.g. "tree forest" that represents a tree as a plant and "tree structure" which represents tree as a mathematical structure.

To get the embedding of the keyword, we simply average all the words in the keyword:

@@

\subsection{Context embedding}
\label{sec:ctx_embed}
\textit{Context} is a sequence of words that contains an ambiguous word within, usually in the middle.
It is parametrized by context length \(l\), which specifies how many of words from both side of the ambiguous word in question were taken into consideration.

@@

Text, that we take context from, must be preprocessed.
We remove any special characters and stopwords, and use lower-case letters only.

\underline{Example with \(l=3\)}:

@@

\subsection{Optimization}
\label{sec:optimization}
As a baseline, we use keyword and context embeddings described above to find the closest keyword given some context, using cosine distance as a metric.
To optimize this system, we use texts from the training set to improve the quality of keyword embeddings.
For each training example, we shift the correct keyword embedding by a factor of \(\alpha\) in the direction of context embedding, which describes said keyword.

@@

@@

@@

\begin{figure}[H]
    \centering
    \caption{Top-1 and top-3 accuracies of the baseline model with different context lengths.}
    \label{fig:baseline}
    \includegraphics[scale=0.7]{res/baseline_context_top_acc.png}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \caption{Comparison of the baseline model (\ref{sec:baseline}) and a model optimized with \(\alpha=0.2\) for different context lengths.}
    \label{fig:exp1_optimization}
    \includegraphics[scale=0.7]{res/exp1_context_top_acc.png}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \caption{Influence of \(\alpha\) parameter value on top-1 and top-3 accuracies for \(l=3\). Blue line denotes the average result from 30 runs, dotted lines denote worst and best result. The value for \(\alpha=0\) is the result achieved by the baseline model (\ref{sec:baseline}).}
    \label{fig:exp2_alpha}
    \includegraphics[scale=0.7]{res/exp2_alpha_top_acc.png}
\end{figure}

@@

\subsection{Beta optimization}
We proved that moving keyword embeddings closer to their contexts is beneficial for the accuracy on the test set (\ref{sec:exp_alpha}).
We theorize, that the results could be further improved, if we also increase the distance between closest incorrect keyword embeddings and the context embedding for a specific example (Eq. \ref{eq:beta_optimization}).

@@

\begin{figure}[H]
    \centering
    \caption{Influence of \(\beta\) parameter value on top-1 and top-3 accuracies for \(l=3\) nad \(\alpha=0.25\). The value for \(\beta=0\) is the best result achieved by the model from previous experiment (Fig. \ref{fig:exp2_alpha}).}
    \label{fig:exp3_beta}
    \includegraphics[scale=0.7]{res/exp3_beta_top_acc.png}
\end{figure}

@@

\subsection{Accuracy in time}
In previous experiments, we passed through the training data only once, which may not be optimal.
The goal of this experiment is to check how a number of epochs (full passes through training data) affects the final accuracy.
We run 5 epochs of training, using best parameters from previous experiments: \(l=3, \alpha=0.25, \beta=0.05\) (Fig. \ref{fig:exp4_1_epochs_fixed}).

\begin{figure}[H]
    \centering
    \caption{Top-1 and top-3 accuracy for model with \(l=3, \alpha=0.25, \beta=0.05\) trained for 5 epochs.}
    \label{fig:exp4_1_epochs_fixed}
    \includegraphics[scale=0.65]{res/exp4_epoch_top_acc.png}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \caption{Comparison of models optimized with different \(\alpha\) and \(\beta\) values with \(l=3\) for 5 epochs.}
    \label{fig:exp4_2_epochs_comparison}
    \includegraphics[scale=0.65]{res/exp4_epoch_multiconfiguration_top_acc.png}
\end{figure}

>>>>

%% Useful packages
%\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{tabu}
\usepackage{array}
\usepackage{tcolorbox}

@@

\section{Our method}
\subsection{Keyword embedding}
\label{sec:kw_embed}
\textit{Keyword} is a sequence of words that lets us disambiguate between different meanings of the same ambiguous word, e.g. \textit{tree (forest)} that represents a tree as a plant and \textit{tree (structure)} which represents tree as a mathematical structure.

To get the embedding of the keyword, we average embeddings of all the words in the keyword:

@@

Example for keyword \textit{tree (forest)}:

\begin{equation}
    \label{eq:example_kw_embed}
    \bm{k}_{tree (forest)} = e(tree, forest) = \frac{e(tree) + e(forest)}{2}
\end{equation}

\subsection{Context embedding}
\label{sec:ctx_embed}
\textit{Context} is a sequence of words that contains an ambiguous word within, usually in the middle.
It is parametrized by \textbf{context length \(l\)}, which specifies how many of words from both side of the ambiguous word in question were taken into consideration.

@@

Text, that we take context from, must be preprocessed.
We remove any special characters and stopwords, and use lower-case letters only. Below is an example taken from the dataset (\ref{sec:dataset}):

\begin{tcolorbox}[
    colframe=blue!25,
    colback=white!10,
    coltitle=white!20!black,
    title={Example for keyword \textit{pool (computer science)} with context length \(l=3\)}]

@@

\textbf{Context embedding}:
\(\bm{c} = \frac{e(kept) + e(ready) + e(use) + e(pool) + e(rather) + e(allocating) + e(destroying)}{7}\)

\end{tcolorbox}

\subsection{Basic model}
\label{sec:basic_model}
The basic approach is to use keyword (\ref{sec:kw_embed}) and context embeddings (\ref{sec:ctx_embed}) to find the closest keyword given some context, using cosine distance as a metric.

In other words, given an input text and marked ambiguous word within, we extract the context and compute its embedding \(\bm{c}\) (see example in the frame above). The keyword, whose embedding is closest to \(\bm{c}\) w.r.t. cosine distance, is chosen as the ambiguous word's meaning.

We evaluate this model based on top-k accuracy, which checks, if the embedding for correct keyword is among closest \(k\) keyword embeddings (\ref{sec:baseline}).

\subsection{Optimization}
\label{sec:optimization}
We assume, that the performance of the basic model (\ref{sec:basic_model}) (and the quality of keyword embeddings) can be improved, if we provide examples of contexts that the specific keyword appears in. To this end, we move keyword embeddings closer to embeddings of contexts they appear in, using examples taken from the training set (\ref{sec:dataset}).

For each training example, we shift the correct keyword embedding by a factor of \(\alpha\), in the direction of the context embedding, which describes said keyword.

@@

Both optimization methods can be seen on Fig. \ref{fig:opt}. The performance of these techniques is evaluated in \ref{sec:exp_alpha} and \ref{sec:exp_beta}.

\begin{figure}
    \centering
    \caption{Optimization. Square is the context embedding \(\bm{c}\), triangles are the closest keyword embeddings, with \(\bm{k}\) being the correct keyword. Without the optimization, the closest keyword embedding w.r.t to context is \(\bm{m}_{1}\), which is incorrect (a). We move the correct keyword embedding in the direction of the context embedding by the factor of \(\alpha\) (Eq. \ref{eq:alpha_optimization}) and the closest \(N\) incorrect keyword embeddings (here \(N=2\)) away from the context embedding by the factor of \(\beta\) (Eq. \ref{eq:beta_optimization}) (b). After optimization, \(\bm{k}\) is the closest keyword (c).}
    \label{fig:opt}
    \includegraphics[scale=0.31]{res/opt.png}
\end{figure}

@@

Finally, we run experiments for a very small number of ambiguous words and meanings (Table \ref{table:dataset}). Our method could have problems with a bigger dataset, since it would be much more difficult to separate different keywords.

@@

\begin{figure}
    \centering
    \caption{Top-1 and top-3 accuracies of the baseline model with different context lengths.}
    \label{fig:baseline}
    \includegraphics[scale=0.7]{res/baseline_context_top_acc.png}
\end{figure}

@@

\begin{figure}
    \centering
    \caption{Comparison of the baseline model (\ref{sec:baseline}) and a model optimized with \(\alpha=0.2\) for different context lengths.}
    \label{fig:exp1_optimization}
    \includegraphics[scale=0.7]{res/exp1_context_top_acc.png}
\end{figure}

@@

\begin{figure}
    \centering
    \caption{Influence of \(\alpha\) parameter value on top-1 and top-3 accuracies for \(l=3\). Blue line denotes the average result from 30 runs, dotted lines denote worst and best result. The value for \(\alpha=0\) is the result achieved by the baseline model (\ref{sec:baseline}).}
    \label{fig:exp2_alpha}
    \includegraphics[scale=0.7]{res/exp2_alpha_top_acc.png}
\end{figure}

@@

\subsection{Beta optimization}
\label{sec:exp_beta}
We proved that moving keyword embeddings closer to their contexts is beneficial for the accuracy on the test set (\ref{sec:exp_alpha}).
We theorize, that the results could be further improved, if we also increase the distance between closest incorrect keyword embeddings and the context embedding for a specific example (Eq. \ref{eq:beta_optimization}).

@@

\begin{figure}
    \centering
    \caption{Influence of \(\beta\) parameter value on top-1 and top-3 accuracies for \(l=3\) nad \(\alpha=0.25\). The value for \(\beta=0\) is the best result achieved by the model from previous experiment (Fig. \ref{fig:exp2_alpha}).}
    \label{fig:exp3_beta}
    \includegraphics[scale=0.7]{res/exp3_beta_top_acc.png}
\end{figure}

@@

\subsection{Accuracy in time}
\label{sec:exp_epochs}
In previous experiments, we passed through the training data only once, which may not be optimal.
The goal of this experiment is to check how a number of epochs (full passes through training data) affects the final accuracy.
We run 5 epochs of training, using best parameters from previous experiments: \(l=3, \alpha=0.25, \beta=0.05\) (Fig. \ref{fig:exp4_1_epochs_fixed}).

\begin{figure}
    \centering
    \caption{Top-1 and top-3 accuracy for model with \(l=3, \alpha=0.25, \beta=0.05\) trained for 5 epochs.}
    \label{fig:exp4_1_epochs_fixed}
    \includegraphics[scale=0.65]{res/exp4_epoch_top_acc.png}
\end{figure}

@@

\begin{figure}
    \centering
    \caption{Comparison of models optimized with different \(\alpha\) and \(\beta\) values with \(l=3\) for 5 epochs.}
    \label{fig:exp4_2_epochs_comparison}
    \includegraphics[scale=0.65]{res/exp4_epoch_multiconfiguration_top_acc.png}
\end{figure}
